---
title: "Introduction to rpbnb"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to rpbnb}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(rpbnb)
d <- read.csv(system.file("extdata", "rwm1984_clean.csv", package = "rpbnb"))
```

## Bivariate NB

Fit a Famoye/Sarmanov bivariate negative binomial model to doctor and hospital
visits:

```{r}
fit <- fit_bnb(docvis ~ outwork, hospvis ~ outwork, data = d,
               dependence = "famoye")
summary(fit)
```

Goodness of fit:

```{r}
g <- bnb_gof(fit, print_output = FALSE)
g$AIC
g$pseudoR2
```

Compare against the independence model (two univariate NB2 margins):

```{r}
fit_ind <- fit_bnb(docvis ~ outwork, hospvis ~ outwork, data = d,
                   dependence = "independence")
c(famoye = as.numeric(logLik(fit)), independence = as.numeric(logLik(fit_ind)))
```

## Marginal effects

```{r}
bnb_marginal_effects(fit, which = "y1", type = "AME", print_output = FALSE)
```

## Random parameters

Let the coefficient on `kids` vary across individuals in the doctor-visits
equation. For a fast vignette we use a small subsample and a modest number of
simulation draws; in practice use the full data with more draws (e.g.
`draws = 1000`) and `compute_se = TRUE`.

```{r}
set.seed(1)
d_small <- d[sample(nrow(d), 600), ]
rp <- fit_rpbnb(docvis ~ outwork + kids, hospvis ~ outwork, data = d_small,
                random_1 = "kids", draws = 100, seed = 1,
                control = rpbnb_control(compute_se = FALSE))
coef(rp)
```

## Simulation

Generate data from a known random-parameter process and recover the parameters:

```{r}
sim <- simulate_rpbnb(n = 1000,
  beta1 = c("(Intercept)" = 0.2, x1 = 0.4),
  beta2 = c("(Intercept)" = 0.1, x1 = -0.3),
  random_1 = list(x1 = list(sd = 0.5)),
  dispersion = c(m1 = 0.4, m2 = 0.5), seed = 1)
head(sim$data)
```
