---
title: "Introduction to rpbnb"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to rpbnb}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(rpbnb)
d <- read.csv(system.file("extdata", "rwm1984_clean.csv", package = "rpbnb"))
```

## Bivariate NB

Fit a Famoye/Sarmanov bivariate negative binomial model to doctor and hospital
visits:

```{r}
fit <- fit_bnb(docvis ~ outwork, hospvis ~ outwork, data = d,
               dependence = "famoye")
summary(fit)
```

Goodness of fit:

```{r}
g <- bnb_gof(fit, print_output = FALSE)
g$AIC
g$pseudoR2
```

Compare against the independence model (two univariate NB2 margins):

```{r}
fit_ind <- fit_bnb(docvis ~ outwork, hospvis ~ outwork, data = d,
                   dependence = "independence")
c(famoye = as.numeric(logLik(fit)), independence = as.numeric(logLik(fit_ind)))
```

## Copula dependence

As an alternative to Famoye/Sarmanov dependence, `fit_bnb()` also supports a
discrete-copula bivariate NB via `dependence = copula(family)`, with `family`
one of `"frank"`, `"normal"` (Gaussian), or `"kimeldorf"` (Clayton). The
dependence parameter is estimated from the data rather than fixed.

```{r}
fit_cop <- fit_bnb(docvis ~ outwork, hospvis ~ outwork, data = d,
                   dependence = copula("normal"))
fit_cop$cop_par   # estimated native parameter (rho, for the Gaussian copula)
fit_cop$cop_tau   # implied Kendall's tau
```

`fit_rpbnb()` accepts the same `dependence = copula(...)` argument to combine
copula dependence with random coefficients (see the copula example at the end
of the next section).

## Marginal effects

```{r}
bnb_marginal_effects(fit, which = "y1", type = "AME", print_output = FALSE)
```

### Interpreting random-parameter fits

For `fit_rpbnb()` models the conditional mean is the population average
`E[exp(x'beta)]` integrated over the random-coefficient distribution, so
marginal effects and elasticities are computed on that integrated mean rather
than on `exp(x'beta)`. Use `rpbnb_marginal_effects()` and `rpbnb_elasticities()`
(the random-parameter analogues of `bnb_marginal_effects()` /
`bnb_elasticities()`):

```r
rpbnb_marginal_effects(fit, which = "both", type = "AME")
rpbnb_elasticities(fit, which = "both", type = "AME")
```

For a continuous covariate the marginal effect averages the per-draw realized
coefficient times `exp(lp)`; for a 0/1 covariate it is the integrated difference
`E[Y | x = 1] - E[Y | x = 0]`. Standard errors come from a numeric delta method
over each equation's mean and log-scale parameters.

## Random parameters

Let the coefficient on `kids` vary across individuals in the doctor-visits
equation. For a fast vignette we use a small subsample and a modest number of
simulation draws; in practice use the full data with more draws (e.g.
`draws = 1000`) and `compute_se = TRUE`.

```{r}
set.seed(1)
d_small <- d[sample(nrow(d), 600), ]
rp <- fit_rpbnb(docvis ~ outwork + kids, hospvis ~ outwork, data = d_small,
                random_1 = "kids", draws = 100, seed = 1,
                control = rpbnb_control(compute_se = FALSE))
coef(rp)
```

`fit_rpbnb()` also accepts `dependence = copula(...)` to pair random
coefficients with copula dependence. Random coefficients are only well
identified on **continuous** regressors under the copula path -- a random
coefficient on a 0/1 dummy trades off against the NB dispersion and is weakly
identified at realistic sample sizes, so we put the random coefficient on the
continuous `age` variable instead of a dummy:

```{r}
rp_cop <- fit_rpbnb(docvis ~ outwork + age, hospvis ~ outwork, data = d_small,
                    random_1 = "age", dependence = copula("normal"),
                    draws = 100, seed = 1,
                    control = rpbnb_control(compute_se = FALSE))
coef(rp_cop)
tanh(coef(rp_cop)[["z_theta"]])  # estimated copula rho
```

## Simulation

Generate data from a known random-parameter process and recover the parameters:

```{r}
sim <- simulate_rpbnb(n = 1000,
  beta1 = c("(Intercept)" = 0.2, x1 = 0.4),
  beta2 = c("(Intercept)" = 0.1, x1 = -0.3),
  random_1 = list(x1 = list(sd = 0.5)),
  dispersion = c(m1 = 0.4, m2 = 0.5), seed = 1)
head(sim$data)
```

`simulate_rpbnb_copula()` does the same, but from a copula-dependent process
(the joint pmf is built from the two NB2 marginal CDFs and a copula CDF,
rather than the Famoye/Sarmanov tilt):

```{r}
sim_cop <- simulate_rpbnb_copula(n = 1000,
  beta1 = c("(Intercept)" = 0.2, x1 = 0.4),
  beta2 = c("(Intercept)" = 0.1, x1 = -0.3),
  random_1 = list(x1 = list(sd = 0.5)),
  dispersion = c(m1 = 0.4, m2 = 0.5),
  copula = copula("normal", par = 0.5), seed = 1)
head(sim_cop$data)
sim_cop$true$tau   # true Kendall's tau implied by rho = 0.5
```
