---
title: "rpbnb Truck Open v2: Coefficient P-values"
---

::: {.callout-note}
This report was generated using AI under general human direction. At the time of generation, the contents have not been comprehensively reviewed by a human analyst.
:::

# Coefficient p-values across scenarios

This table compares two-sided asymptotic Wald p-values for the regression coefficients across six TMB fits. The model and data were held fixed while estimation method and copula family varied. Values are rounded to three decimal places; values smaller than 0.0005 display as `0.000`.

```{r load-data}
#| label: load-data
library(knitr)

pvalue_file <- file.path("..", "results", "benchmark_open_v2_significance.rds")
stopifnot(file.exists(pvalue_file))
sig <- readRDS(pvalue_file)

pvalues <- sig$pwide
names(pvalues)[names(pvalues) == "parameter"] <- "Coefficient"

# Rename scenario columns for readability.
scenario_names <- c(
  p.sml_frank = "SML + Frank",
  p.sml_kimeldorf = "SML + Kimeldorf",
  p.sml_normal = "SML + Normal",
  p.laplace_frank = "Laplace + Frank",
  p.laplace_kimeldorf = "Laplace + Kimeldorf",
  p.laplace_normal = "Laplace + Normal"
)
names(pvalues) <- ifelse(names(pvalues) %in% names(scenario_names),
                          scenario_names[names(pvalues)], names(pvalues))

# Display exactly three decimal places, including trailing zeroes.
pvalues[-1] <- lapply(pvalues[-1], function(x) formatC(x, format = "f", digits = 3))
```

```{r pvalue-table}
#| label: pvalue-table
kable(pvalues, align = "lrrrrrr", caption = "Two-sided coefficient p-values by estimation method and copula family.")
```

The p-values are unadjusted for multiple comparisons and use the standard 0.05 significance threshold.
